---
title: "PubH_7462_HW2_Shen0464"
author: "You Shan Shen"
date: "2022/2/6"
output: github_document
---

```{r Setup }
library(knitr)
knitr::opts_knit$set(echo = TRUE, warning = FALSE, results = "hide", error = FALSE, message = FALSE, include = FALSE,
root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)

theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "right"))
```

3.1 Data Exploration & Cleaning (10pts)

```{r pressure, echo=FALSE}
library(dplyr)
library(tidyr)
library(stringr)

brfss_smart_df <- 
  read.csv("./data/brfss_smart_2010.csv") %>%
  as_tibble() %>% 
  janitor::clean_names() %>% 
  filter(topic %in% 'Overall Health') %>%
  dplyr::select(
  year, response, locationdesc, sample_size, data_value
  ) %>%
  separate(locationdesc, sep = " - ", into = c("state", "county")) %>%
  rename( 
    'prop. of responses' = data_value, 
    'sample size'= sample_size) %>%
  mutate( year = as.numeric(year),
          state = factor(state),
         response = as.factor(response),
         county= as.character(county)) 

```



3.2 Data Description (10pts)
```{r}
number_row <- brfss_smart_df %>% nrow()
number_col <- brfss_smart_df %>% ncol()
year_level <- length(unique(brfss_smart_df$year))
response_level <- length(unique(brfss_smart_df$response))
number_state <- length(unique(brfss_smart_df$state))
number_county <- length(unique(brfss_smart_df$county))
```
 
2002-2010. BRFSS SMART County Prevalence contains `r number_row` observations and `r number_col` variables. Each observation is defined as a data about a U.S. resident regarding their health-related risk behaviors, chronic health conditions, and use of preventive services.

The definition of variables in 3.1:

Year: The survey conduct in `r year_level` different numbers of year.

Response: There are `r response_level` levels response from an observation.

State: Observations are from `r number_state` different states in US.

County: Observations are from `r number_state` different counties in US.

Sample size: the number of sample from a specific year, response, state, and county.

Proportion of response: the proportion of sample from a specific level of response according to year, state, and county in percentage.



3.3 Do Data Science (50pts)
3.3.1 In the year 2004, which states were observed at 6 locations? (10pts)
```{r}
devtools::install_github("rstudio/gt")

library(gt)
library(webshot)

  brfss_smart_df %>% 
  filter(year %in% '2004') %>% 
  dplyr::select( state, county ) %>%
  group_by( state) %>% 
  distinct( county , .keep_all = TRUE) %>%
  summarize(n = n() ) %>%
  filter(n == 6) %>%
  gt() 
  

```



3.3.2 Make a “spaghetti plot” that shows the number of observed locations in each state from
2002 to 2010. Which state has the highest mean number of locations over this period? (10pts)

```{r}
library(ggplot2)
library(viridis)
library(forcats)

brfss_smart_spa <- brfss_smart_df %>%
  group_by(state, year) %>%
  distinct(state, year, county) %>%
  summarize(number = n() ) 

brfss_smart_spa$state1 <- fct_reorder(brfss_smart_spa$state,brfss_smart_spa$number,mean ,  na.rm = TRUE, .desc = TRUE)

brfss_smart_spaggplot <- brfss_smart_spa  %>%
   ggplot(aes(x = year, y = number, colour = state1)) +
   geom_line(aes(group = state)) +
   scale_color_viridis_d( "State" )

brfss_smart_spaggplot
ggsave("./figures/brfss_smart_spaggplot.png", brfss_smart_spaggplot)
```



3.3.3 Make a table showing, for the years 2002, 2006, and 2010, the mean and standard
deviation of sample size and proportion of Excellent, Good, and Poor responses across locations
in MN. (15pts)

```{r}
library(gt)
MN_brfss_smart <- 
  brfss_smart_df %>%
  filter(state %in% 'MN') %>%
  filter(year %in% c('2002', '2006', '2010')) %>%
  filter(response %in% c('Excellent', 'Good', 'Poor')) %>%
  dplyr::select( year, response, `sample size`, `prop. of responses` )


  summary_MN_brfss_smart <-
  MN_brfss_smart %>%
  group_by(year, response) %>%
  summarise( across(c(`sample size`, `prop. of responses`),
                    list(mean = mean, sd = sd),
                    na.rm = TRUE, 
                    .names = "{.col}_{.fn}"))
   
  summary_MN_brfss_smart %>%
    mutate(across(where(is.numeric), round, 3)) %>%
   gt() %>%
  tab_header("Mean and SD of Sample Size, Prop. Response from 2002, 2006,and 2010 in MN ") 
 

```

3.3.4 Create a ggplot that communicates the results/trends from the table above and stands
on its own (15pts)

```{r}
library(stringr)

tidy_brfss_smart <-
summary_MN_brfss_smart %>%
pivot_longer(cols = starts_with("sample"),
names_to = "sample_size_type", 
values_to = "sample_values") %>%
mutate(
sample_size_type = stringr::str_remove(sample_size_type, "sample size_") %>%
str_to_upper() %>% 
as.factor()
) %>%
pivot_longer(cols = starts_with("prop."), 
names_to = "prop._type",
values_to = "prop_values") %>%
mutate(
prop._type = stringr::str_remove(prop._type, "prop. of responses_") %>%
str_to_upper() %>% 
as.factor())



base.gg <- 
  tidy_brfss_smart %>%
  mutate(year= as.factor(year)) 

year_ggplot <-  base.gg %>%
ggplot(aes(x = year, y = `sample_values`, colour = response, fill = response))+
geom_col()  +
ggtitle( "Mean of Sample Size for Response from 2002, 2006,and 2010 in MN" ) +
xlab( "Year" ) + 
ylab( "Mean of Sample Size Values" )

year_ggplot

response_type_ggplot <-
  year_ggplot  + 
  facet_wrap(~ response, scales = "free_x") +
ggtitle( "Mean of Sample Size for Response from 2002, 2006,and 2010 in MN" ) +
xlab( "Year" ) + 
ylab( " Mean of Sample Size Values" )

response_type_ggplot

dir.create("./figures", showWarnings = FALSE)
ggsave("./figures/year_ggplot.png", year_ggplot)
ggsave("./figures/response_type_ggplot.png", response_type_ggplot)

```